2H qSIP Calculations

### 2021-10-22 qSIP_calculations.Rmd
many more plot types, FID data now integrated

## Read raw data files

@@

# no zero samples
peaks_mapped_with_zeros %>% filter(is.na(zero_sample_id)) %>% View()
```

@@

```{r}
p_gen_time_dist <- gen_calc_data %>% filter(inc_time_d != 0) %>% 
  ggplot(
  aes(x = gen_d, 
      y = inc_time_d_str, 
      color = inc_time_d_str, 
      fill = inc_time_d_str)) +
  stat_slab(height = 0.3, alpha = 0.7) +
  # stat_pointinterval(color = "black") +
  geom_boxplot(fill = "white",
               color = "black",
               width = 0.1,
               outlier.shape = NA) +
  geom_point(
    color = "black",
    #shape = 124,
    #size = 3,
    alpha = 0.5,
    position = position_nudge(y = -.03)) +
  # geom_text_repel(aes(label = compound), 
  #                 color = "black", 
  #                 nudge_y = -0.3, 
  #                 size = 3) +
  scale_x_log10() +
  annotation_logticks(sides = "b") +
  scale_fill_npg() +
  scale_color_npg() +
  theme_classic() +
  labs(x = "Generation time (days)",
       y = "Incubation time",
       title = "Generation times of soil bacteria",
       subtitle = "Lipid-specific generation times of bacteria collected from Alpine Tundra",
       caption = "Soils collected from Colorado") +
  theme(legend.position = "none",
        axis.ticks.y = element_blank())
p_gen_time_dist

@@

@@

# GC-FID Data

all_data <-
  tibble(
    sample = sample_names,
    file = file_list,
    data = map(file.path(root, file), read_chromeleon_export),
    injection_details = map(data, "injection_details"),
    integration_results = map(data, "integration_results")
  ) %>% select(-data)
```

## Unnest integration details

```{r}
results <- all_data %>% 
  select(-injection_details) %>% 
  unnest(integration_results) %>% 
  filter(nchar(`Peak Name`) > 0) %>% 
  # major compound groups
  mutate(
    group = `Peak Name` %>% str_remove("\\d+-OH") %>% parse_number() %>% abs() %>% {paste0("C", .)} %>% factor()
  )
```

```{r}
# Combine triplicate samples into one
results_trip <- results %>% 
  # Remove standards and fractions
  filter(sample != "FAME37.1") %>% 
  filter(sample != "TCBAME-2") %>% 
  filter(!str_detect(sample, "F1")) %>% 
  filter(!str_detect(sample, "F2")) %>% 
  filter(!str_detect(sample, "F3")) %>% 
  filter(!str_detect(sample, "cr")) %>%
  mutate(sample_rep = str_remove(sample, "[xyz]")) %>% 
  group_by(sample_rep, `Peak Name`) %>%
  # Generate summary stats
  summarize(rel_area_max = max(`Relative Area`, na.rm=TRUE),
            rel_area_min = min(`Relative Area`, na.rm=TRUE),
            rel_height_max = max(`Relative Height`, na.rm=TRUE),
            rel_height_min = min(`Relative Height`, na.rm=TRUE),
            `Retention Time` = mean(`Retention Time`, na.rm=TRUE),
            `Relative Area` = mean(`Relative Area`, na.rm=TRUE),
            `Relative Height` = mean(`Relative Height`, na.rm=TRUE)) %>% 
  # Add some basic metadata
  mutate(inc_time = case_when(
    str_detect(sample_rep, "...0") ~ 0,
    str_detect(sample_rep, "...3") ~ 3,
    str_detect(sample_rep, "...7") ~ 7),
    site = case_when(
      str_detect(sample_rep, "c") ~ "Conifer",
      str_detect(sample_rep, "t") ~ "Tundra",
      str_detect(sample_rep, "g") ~ "Grassland",
      str_detect(sample_rep, "m") ~ "Meadow")
  )
```

## PLFA Summary Plot
```{r}
p_72hr_FAs <- results_trip %>% filter(inc_time == "3") %>% 
  # Remove Standard Peaks
  #filter(str_detect(`Peak Name`, c("PAIBE", "21:0 (STD)", "23:0 (STD)"), negate = TRUE)) %>% 
  filter(`Peak Name` %nin% c("PAIBE", "21:0 (STD)", "23:0 (STD)")) %>% 
  ggplot(aes(x = `Retention Time`, 
             y = `Relative Area`, 
             color = sample_rep, 
             fill = sample_rep, 
             label = `Peak Name`)) +
  geom_col(width = 0.08) +
  geom_errorbar(aes(ymin=rel_area_min, 
                    ymax = rel_area_max), 
                color = "black", 
                width = 0.05) +
  facet_wrap(~sample_rep, 
             ncol = 1, 
             scales = "free_y") +
  geom_label_repel(fill = "black", 
                   color = "white", 
                   alpha = 0.7, 
                   size = 3, 
                   nudge_y = 2, 
                   segment.color = "black") +
  ggtitle("PLFA Profile: 72hr") +
  theme_bw() +
  scale_color_jama() +
  scale_fill_jama()
p_72hr_FAs
```

### IRMS relabund
```{r}
peaks_mapped_active %>% ggplot(aes(x = rt, y = rel_amount, fill = inc_time_d_str, color = inc_time_d_str)) +
  geom_col() +
  facet_wrap(~inc_time_d_str, ncol = 1)
```

### Relabund vs. activity
```{r}
gen_calc_data %>%
  filter(inc_time_d != 0) %>% 
  ggplot(
  aes(x = rel_amount, 
      y = u_d, 
      color = inc_time_d_str, 
      label = compound)) +
  geom_point() +
  geom_smooth(method = "lm",
              se = FALSE,
              formula = "y~x") +
  geom_line(aes(group = compound), 
            color = "black", 
            alpha = 0.3) +
  geom_text_repel(size = 3) +
  theme_classic() +
  scale_color_npg() +
  labs(x = "Relative Amount",
       y = "Turnover (d^-1)",
       color = "Incubation Time") +
  theme(legend.position = "top")
```

### Relabund vs. activity colored by 2F

```{r}
gen_calc_data %>%
  filter(inc_time_d != 0) %>% 
  ggplot(
  aes(x = rel_amount, 
      y = u_d, 
      color = as.numeric(f_t), 
      label = compound)) +
  geom_point(aes(shape = inc_time_d_str)) +
  geom_smooth(method = "lm",
              se = FALSE,
              formula = "y~x") +
  geom_line(aes(group = compound), 
            color = "black", 
            alpha = 0.3) +
  geom_text_repel(color = "black", 
                  size = 3) +
  scale_color_viridis_c(option = "magma") +
  theme_classic() +
  labs(x = "Relative Amount",
       y = "Turnover (d^-1)",
       shape = "Incubation Time",
       color = "2F") +
  theme(legend.position = "bottom")
```

## Merge IRMS and FID data

### PICK UP HERE: how do we match GC-FID compound data with IRMS compound data?

```{r}
results_trip_matched <- results_trip %>% 
  filter(`Peak Name` %in% gen_calc_data$compound)

matched_idx <- match(gen_calc_data$compound, results_trip_matched$`Peak Name`)
results_trip_matched$`Peak Name`[matched_idx]

combined_irms_fid <- gen_calc_data %>%
  group_by(terrain) %>% 
  mutate(relative_area = results_trip$`Relative Area`[""])
```

>>>>

FID_all_data <-
  tibble(
    sample = sample_names,
    file = file_list,
    data = map(file.path(root, file), read_chromeleon_export),
    injection_details = map(data, "injection_details"),
    integration_results = map(data, "integration_results")
  ) %>% select(-data)
```

### Unnest integration details

```{r}
FID_results <- FID_all_data %>% 
  select(-injection_details) %>% 
  unnest(integration_results) %>% 
  filter(nchar(`Peak Name`) > 0) %>% 
  # Specify chain length
  mutate(
    chain = `Peak Name` %>% 
      str_remove("\\d+-OH") %>% 
      parse_number() %>% 
      abs() %>% 
      {paste0("C", .)} %>% 
      factor()
      ) %>% 
  # Rename these columns so we can join the IRMS and FID data
  # Important to specify what data comes from FID!
  rename("sample_id" = sample,
         "compound" = `Peak Name`,
         "rt_FID" = `Retention Time`,
         "area_FID" = `Area`,
         "height_FID" = `Height`,
         "rel_area_FID" = `Relative Area`,
         "rel_height_FID" = `Relative Height`,
         "amount_FID" = `Amount`
         )
```

## Read raw IRMS data files

@@

# no zero samples
peaks_mapped_with_zeros %>% filter(is.na(zero_sample_id)) %>% View()

peaks_mapped_with_zeros <- peaks_mapped_with_zeros %>% 
  mutate(
    sample_id = substr(file_id, 10,14)
    )
```

## Join IRMS and GC-FID Data

```{r}
peaks_mapped_with_zeros_FID <- peaks_mapped_with_zeros %>% 
  left_join(FID_results, by = c("sample_id", "compound"))
```

```{r}
# Combine triplicate samples into one
FID_results_trip <- FID_results %>% 
  # Remove standards and fractions
  filter(sample != "FAME37.1") %>% 
  filter(sample != "TCBAME-2") %>% 
  filter(!str_detect(sample, "F1")) %>% 
  filter(!str_detect(sample, "F2")) %>% 
  filter(!str_detect(sample, "F3")) %>% 
  filter(!str_detect(sample, "cr")) %>%
  mutate(sample_rep = str_remove(sample, "[xyz]")) %>% 
  group_by(sample_rep, `Peak Name`) %>%
  # Generate summary stats
  summarize(rel_area_max = max(`Relative Area`, na.rm=TRUE),
            rel_area_min = min(`Relative Area`, na.rm=TRUE),
            rel_height_max = max(`Relative Height`, na.rm=TRUE),
            rel_height_min = min(`Relative Height`, na.rm=TRUE),
            `Retention Time` = mean(`Retention Time`, na.rm=TRUE),
            `Relative Area` = mean(`Relative Area`, na.rm=TRUE),
            `Relative Height` = mean(`Relative Height`, na.rm=TRUE)) %>% 
  # Add some basic metadata
  mutate(inc_time = case_when(
    str_detect(sample_rep, "...0") ~ 0,
    str_detect(sample_rep, "...3") ~ 3,
    str_detect(sample_rep, "...7") ~ 7),
    site = case_when(
      str_detect(sample_rep, "c") ~ "Conifer",
      str_detect(sample_rep, "t") ~ "Tundra",
      str_detect(sample_rep, "g") ~ "Grassland",
      str_detect(sample_rep, "m") ~ "Meadow")
  )
```

@@

```{r}
p_gen_time_dist <- gen_calc_data %>% filter(inc_time_d != 0) %>% 
  ggplot(
  aes(x = gen_d, 
      y = inc_time_d_str, 
      color = inc_time_d_str, 
      fill = inc_time_d_str)) +
  stat_slab(height = 0.3, alpha = 0.7) +
  # stat_pointinterval(color = "black") +
  geom_boxplot(fill = "white",
               color = "black",
               width = 0.1,
               outlier.shape = NA) +
  geom_point(
    color = "black",
    #shape = 124,
    #size = 3,
    alpha = 0.5,
    position = position_nudge(y = -.03)) +
  # geom_text_repel(aes(label = compound), 
  #                 color = "black", 
  #                 nudge_y = -0.3, 
  #                 size = 3) +
  scale_x_log10() +
  annotation_logticks(sides = "b") +
  scale_fill_npg() +
  scale_color_npg() +
  theme_classic() +
  labs(x = "Generation time (days)",
       y = "Incubation time",
       title = "Generation times of soil bacteria",
       subtitle = "Lipid-specific generation times of bacteria",
       caption = "Soils collected from Colorado") +
  theme(legend.position = "none",
        axis.ticks.y = element_blank())
p_gen_time_dist

@@

## Generation Time, all points, not separated by inc_time

```{r}
p_gen_time_dist_all <- gen_calc_data %>% filter(inc_time_d != 0) %>% 
  ggplot(
  aes(x = gen_d, 
      y = 1)) +
  stat_halfeye(height = 0.1, 
            alpha = 0.7,
            color = "black") +
  geom_jitter(aes(y = 0.9, 
                 color = inc_time_d_str),
             alpha = 0.5,
             height = 0.03,
             size = 1) +
  # geom_text_repel(aes(label = compound), 
  #                 color = "black", 
  #                 nudge_y = -0.3, 
  #                 size = 3) +
  scale_x_log10() +
  coord_cartesian(ylim = c(0.5, 1.25)) +
  annotation_logticks(sides = "b") +
  scale_fill_npg() +
  scale_color_npg() +
  theme_classic() +
  labs(x = "Generation time (days)",
       title = "Generation times of soil bacteria",
       subtitle = "Lipid-specific generation times of bacteria",
       caption = "Soils collected from Colorado",
       color = "Incubation Time") +
  theme(legend.position = "bottom",
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        axis.title.y = element_blank())
p_gen_time_dist_all
```

## Generation time of all points, not log scale

```{r}
p_gen_time_dist_all_lin <- gen_calc_data %>% 
  filter(inc_time_d != 0,
         gen_d >= 0) %>% 
  ggplot(
  aes(x = u_d, 
      y = 1)) +
  stat_halfeye(height = 0.25, 
            alpha = 0.7,
            color = "black") +
  geom_jitter(aes(y = 0.9, 
                 color = inc_time_d_str),
             alpha = 0.5,
             height = 0.03,
             size = 1) +
  # geom_text_repel(aes(label = compound), 
  #                 color = "black", 
  #                 nudge_y = -0.3, 
  #                 size = 3) +
  coord_cartesian(ylim = c(0.8, 1.25)) +
  scale_fill_npg() +
  scale_color_npg() +
  theme_classic() +
  labs(x = "Turnover Time (d^-1)",
       title = "Turnover Times",
       subtitle = "Lipid-specific turnover times of bacteria",
       caption = "Soils collected from Colorado",
       color = "Incubation Time") +
  theme(legend.position = "bottom",
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        axis.title.y = element_blank())
p_gen_time_dist_all_lin
```

@@

```{r}
# coming soon!
```

## Compound Abundance Versus Turnover Time
```{r}
# coming soon!
```
